---
title: "Carolinian Dictionary Quality Checks"
output: html_document
---

```{r setup}
library(tidyverse)

df <- read_tsv("../outputs/carolinian_combined.tsv", col_types = cols(.default = "c")) |>
  mutate(across(everything(), \(x) replace_na(x, "")))

cat("Loaded", nrow(df), "rows\n")
head(df, 3)
```

## Check 1: Headword is not empty

```{r check1}
flagged <- df |> filter(str_trim(Headword) == "")
cat("Flagged:", nrow(flagged), "rows with empty headword\n")
if (nrow(flagged) > 0) print(flagged |> select(Source_Page, Entry_Type, Headword, Gloss))
```

## Check 2: Gloss is not empty (allow if Variant_Form is set)

```{r check2}
flagged <- df |>
  filter(str_trim(Gloss) == "", str_trim(Variant_Form) == "")
cat("Flagged:", nrow(flagged), "rows with empty gloss and no variant form\n")
if (nrow(flagged) > 0) print(flagged |> select(Source_Page, Entry_Type, Headword, Gloss, Variant_Form, Cross_References))
```
Headword: filooris --> Gloss: flower POS: n
Headword: túlungasch --> Gloss: Lips
Others are variants, so OK --> maybe we should exclude them because they should be already captured in Variant_Form

## Check 3: Subentry must have a non-empty Parent_Lexeme

```{r check3}
flagged <- df |>
  filter(Entry_Type == "subentry", str_trim(Parent_Lexeme) == "")
cat("Flagged:", nrow(flagged), "subentries with empty Parent_Lexeme\n")
if (nrow(flagged) > 0) print(flagged |> select(Source_Page, Headword, Parent_Lexeme, Gloss))
```

## Check 4: Duplicate headwords (within same Entry_Type) must have Sense_Number or Homonym_Number

```{r check4}
flagged <- df |>
  group_by(Headword, Entry_Type) |>
  filter(n() > 1) |>
  ungroup() |>
  filter(str_trim(Sense_Number) == "", str_trim(Homonym_Number) == "")
cat("Flagged:", nrow(flagged), "duplicate headwords (within same Entry_Type) missing both Sense_Number and Homonym_Number\n")
if (nrow(flagged) > 0) print(flagged |> select(Source_Page, Entry_Type, Headword, Sense_Number, Homonym_Number, Gloss))
```
Page 136 -- complete hallucination at Stage 2
Index 3993 --> Headword: limal (not mal)
Some are genuine -- the dictionary indeed has these duplicate headwords, but no homonyms indicated
Some are just duplicates from cross-page contintuation (I guess)
falafal --> subentry homonym index not picked up

## Check 5: POS is not empty and uses only known tags

```{r check5}
VALID_POS <- c(
  "adv", "asp", "caus", "cl", "conj", "dem", "dir suf", "idiom", "intj",
  "n", "neg", "num", "num cl", "obj suf", "ord pref", "poss cl", "poss suf",
  "pref", "pron", "ques", "redup", "rel n", "subj", "suf", "tns-asp",
  "v", "vi", "vt",
  "vi caus", "vt caus", "n caus", "poss pron", "n pref",
  "vi (pass)", "vi (redup)", "vi phr", "adv suf", "caus pref", "conj, comp", "dem pron", "dem pron ques", 
  "dem ques", "foc", "intj neg", "loc", "loc vt",
  "n (idiom)", "n (idiom?)", "n (pass)", "n (redup)", "n phr", "n ques", 
  # need to check these cases 
  "n, vi", "n, vi caus", "vi, n", "n caus, vi caus", "n, v","v, n","vi caus, n","vi n","vi, n caus",
  "neg asp", "neg tnsasp", "numcl", 
  "ordinal pref", # this has to be ord pref
  "poss","poss cl, n","Poss suf","prep","pron + asp","pron + neg","pron + tns-asp","prt","ques n","ques pron","v phr","v pref","v suf",
  "vi (idiom)", "vi (prog)", "vi caus (redup)", "vi caus (pass)", "vi caus, redup", "vi, vt", "vt (idiom)", "vt (redup)", "vt (reflex)"
)

# Exclude variant/redirect entries
df_pos <- df |>
  filter(
    !str_starts(str_trim(Gloss), "Variant of"),
    !(str_trim(Variant_Form) != "" & str_trim(Gloss) == "")
  )

# Empty POS
empty_pos <- df_pos |> filter(str_trim(POS) == "")
cat("Flagged:", nrow(empty_pos), "rows with empty POS (excluding variant/redirect entries)\n")
if (nrow(empty_pos) > 0) print(head(empty_pos |> select(Source_Page, Entry_Type, Headword, POS, Gloss), 20))

# Unknown POS
unknown_pos <- df_pos |>
  filter(str_trim(POS) != "", !str_trim(POS) %in% VALID_POS)
cat("\nFlagged:", nrow(unknown_pos), "rows with unrecognised POS\n")
if (nrow(unknown_pos) > 0) {
  print(unknown_pos |> select(Source_Page, Entry_Type, Headword, POS) |> distinct(POS, .keep_all = TRUE))
  cat("Unique unrecognised POS values:", paste(sort(unique(str_trim(unknown_pos$POS))), collapse = ", "), "\n")
}
```
Some POS tags can be incorrectly written (not according to the guideline page)
Need to take care of these dialectal variants as in (TAN), (S) in POS columns
Need to take care of these variants --> maybe just exclude them

```{r}
empty_pos_gloss <- empty_pos %>%
  filter(Gloss != "")
```

Headword: lighitaal36 --> lighitaaló (maybe we can quickly check if headwords have any numbers or other unrelated characters)
Fix subentry headwords with — 
Some POS tags were actually missing from the dictionary
Some haven't picked up because of multiple senses

## Check 6: If Examples is not empty, Example_Glosses must not be empty

```{r check6}
flagged <- df |>
  filter(str_trim(Examples) != "", str_trim(Example_Glosses) == "")
cat("Flagged:", nrow(flagged), "rows with examples but no example gloss\n")
if (nrow(flagged) > 0) print(flagged |> select(Source_Page, Headword, Examples, Example_Glosses))
```

## Polishing

### Deduplication: remove cross-page duplicate entries

```{r dedup_preview}
KEY_COLS <- c("Headword", "Entry_Type", "Parent_Lexeme", "Sense_Number", "Homonym_Number")

df_sorted <- df |>
  mutate(.page_num = as.integer(str_extract(Source_Page, "\\d+"))) |>
  arrange(.page_num)

to_remove <- df_sorted |>
  group_by(across(all_of(KEY_COLS))) |>
  filter(n() > 1) |>
  slice(-1) |>
  ungroup()

cat("Found", nrow(to_remove), "cross-page duplicate rows to remove:\n")
if (nrow(to_remove) > 0) print(to_remove |> select(Source_Page, Entry_Type, Headword, Sense_Number, Homonym_Number, Gloss))
```

```{r dedup_apply}
df_clean <- df_sorted |>
  group_by(across(all_of(KEY_COLS))) |>
  slice(1) |>
  ungroup() |>
  select(-.page_num)

cat("Rows before:", nrow(df), "→ after:", nrow(df_clean), "\n")
```

### Resolve variant entries

```{r variant_preview}
COPY_COLS <- c("Gloss", "Gloss_Secondary", "POS", "Usage_Note", "Phonetic",
               "Cross_References", "Examples", "Example_Glosses",
               "Morphemic_Representation", "Etymology", "Inflection", "Dialect")

is_variant <- str_starts(str_trim(df_clean$Gloss), "Variant of") |
              (str_trim(df_clean$Variant_Form) != "" & str_trim(df_clean$Gloss) == "")

df_variants <- df_clean[is_variant, ] |>
  mutate(.target = coalesce(
    str_trim(str_extract(Gloss, "(?<=Variant of ).+")),
    if_else(str_trim(Variant_Form) != "" & str_trim(Gloss) == "", str_trim(Variant_Form), NA_character_)
  ))

# Build lookup from main entries
lookup <- df_clean |>
  filter(Entry_Type == "main") |>
  distinct(Headword, .keep_all = TRUE)

resolvable <- df_variants |> filter(.target %in% lookup$Headword)
not_found  <- df_variants |> filter(!.target %in% lookup$Headword, !is.na(.target))

cat("Resolvable:", nrow(resolvable), "| Target not found:", nrow(not_found), "\n")
if (nrow(not_found) > 0) {
  cat("Could not resolve (target not yet processed):\n")
  not_found |> select(Headword, .target, Source_Page) |> print()
}
```

```{r variant_apply}
df_resolved <- df_clean

for (i in seq_len(nrow(resolvable))) {
  idx <- which(
    df_resolved$Headword    == resolvable$Headword[i] &
    df_resolved$Entry_Type  == resolvable$Entry_Type[i] &
    df_resolved$Source_Page == resolvable$Source_Page[i]
  )
  if (length(idx) == 0) next
  target_row <- lookup |> filter(Headword == resolvable$.target[i])
  if (nrow(target_row) == 0) next
  for (col in COPY_COLS) {
    if (str_trim(df_resolved[[col]][idx]) == "") {
      df_resolved[[col]][idx] <- target_row[[col]][1]
    }
  }
}

df_resolved <- df_resolved |> select(-.target, -any_of(".target"))
cat("Resolved", nrow(resolvable), "variant entries\n")

write_tsv(df_resolved, "../outputs/carolinian_polished.tsv")
cat("Saved → outputs/carolinian_polished.tsv\n")
```
